---
title: "Sustainable Energy for All - Markdown Report"
author: "Sam Gianan"
date: "2026-07-31"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "hide")
```

# Executive Summary

I'll be dealing with the two main questions from the Tidy Tuesday prompt:

1. Which countries have the lowest capacity for solar energy?
2. What form of renewable energy has, on average, experienced the fastest rate of adoption?

## Loading and Profiling the Data

I first loaded the needed libraries, as well as the .csv file for analysis.

``` {r}
library(tidyverse)
library(ggnewscale)
library(naniar)
library(mice)
library(scales)
library(countrycode)
library(maps)

energy <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-05-26/energy_cleaned.csv")
head(energy)
```

Before I pursue exploratory data analysis, I'll first explain what the variables are; it is vital to first comprehend them before their relationshios.

**country_name**, **country_code**, and **year** are self-explanatory, so let's move on to the remaining 49 variables. Some are already explained in the codebook of the original GitHub repository, so notes here will be about variables that need further explanation. Examples of such variables are component indices, purchasing power parity, primary energy savings, and more.

### Understanding LMDI Analysis

Logarithmic Mean Divisia Index (LMDI) decomposition analysis is a mathematical method that splits the reasons for a change into various factors. An example would be for when one's water bill increases. There may be a variety of factors for this change, such as a systemic increase from the company, an increase in personal water usage, or others. LMDI analysis allows for the full change to be split into these various factors. It's often applied to environmental data and heavily preferred due to fully accounting for changes, leaving no residuals.

This is important to explain three of our variables, **divisia_decomp_analysis_activity_component_index**, **divisia_decomp_analysis_energy_intensity_component_index**, and **divisia_decomp_analysis_structure_component_index**. 

- The activity component index is the change in energy use due to population and economy growth.
- The energy intensity component index measures efficiency; it reflects change in energy use per unit of activity and measures technological efficiency.
- The structure component index showcases the change in energy use due to shifts of the impact shared by various sectors (such as agriculture, manufacturing, etc.).

From the values of the indices, it can be observed that the indices are multiplicative. This means:

- Value > 1 meeans that the factor increased energy usage.
- Value < 1 means that the factor decreased energy usage.
- Value = 1 means that the factor had no effect on energy usage.

### Variable Notes

Here, I'll list brief descriptions and notes on other variables that may not be self-explanatory. These values are similarly described in the codebook, but further information is written here to supply the economic and environmental jargon. Note that for 2005 US dollar variables, we use this value to remove the impact of inflation in calculations.

- **energy_intensity_level_final_energy_megajoules_per_usd_2005_ppp** | This measures how much final energy usage in megajoules is needed to produce 1 US dollar (2005 value) of GDP. *Final energy* refers to the processed energy source, such as electricity, gasoline, etc. 
- **energy_intensity_level_primary_energy_megajoules_per_usd_2005_ppp** | This measures how much primary energy usage in megajoules is needed to produce 1 US dollar (2005 value) of GDP. *Primary energy* refers to the raw energy source, such as sunlight, crude oil, etc.
- **energy_intensity_industrial_sector_megajoules_per_usd_2005** | This measures the energy usage by just the industrial sector, encompassing manufacturing, construction, and production.
- **energy_intensity_other_sectors_megajoules_per_usd_2005** | This measures the energy usage by every other sector *except* the industrial sector.
- **energy_savings_primary_energy_terajoules** | This describes the amount of raw energy sources saved by optimizing processing.
- **final_to_primary_energy_ratio_pct** | This describes how much energy is lost during the extraction, processing, and transmission stages.
- **renewable_energy_installed_capacity_gigawatts**, **share_of_renewable_capacity_in_total_capacity_pct**, and **total_installed_generation_capacity_gigawatts** | *Installed capacity* refers to the maximum total capacity that the power source can produce.
- **thermal_efficiency_in_power_supply_pct** | *Thermal efficiency* measures how well a system can convert heat into power or electricity.

Now that we have a more technical understanding of the dataset's background, we can now move on to exploratory data analysis.

## Data Cleaning and Exploratory Data Analysis

Let's first check our dataset for the countries listed. 

``` {r}
energy %>% 
  distinct(country_name) %>% 
  pull(country_name)
```

Some values in the **country_name** variable refer to region and income. We must then remove these rows. Let us also replace the **yr** variable with **year** as this would be more intuitive. This also applies to **country_name**, becoming **country**.

``` {r}
names_to_remove <- c("Caucasus and Central Asia", "Eastern Asia (including Japan)", "Eastern Asia (not including Japan)", "Eastern Europe", "Europe", "High income", "High income: nonOECD", "High income: OECD", "Latin America and Caribbean", "Low & middle income", "Low income", "Lower middle income", "Middle income", "Northern Africa" , "Nothern America", "Oceania", "Oceania (not including Australia and New Zealand)", "South Eastern Asia", "Southern Asia", "Sub-Saharan Africa", "Upper middle income", "Western Asia", "Western Sahara", "World")

energy <- energy %>%
  filter(!country_name %in% names_to_remove)
energy <- energy %>% rename(year = yr)
energy <- energy %>% rename(country = country_name)
```

To make plotting easier during exploratory data analysis, we'll also need to rename longer variables. Comments are added to explain the categoriazation of variables.

``` {r}
energy <- energy %>%
  rename(
    # access
    rural_access_fuel = access_non_solid_fuel_rural_pop_pct,
    total_access_fuel = access_non_solid_fuel_total_pop_pct,
    urban_access_fuel = access_non_solid_fuel_urban_pop_pct,
    rural_access_elec = access_electricity_rural_pop_pct,
    total_access_elec = access_electricity_total_pop_pct,
    urban_access_elec = access_electricity_urban_pop_pct,
    # indices
    activity_index = divisia_decomp_analysis_activity_component_index,
    efficiency_index = divisia_decomp_analysis_energy_intensity_component_index,
    structure_index = divisia_decomp_analysis_structure_component_index,
    # economics
    final_energy_ppp_mj = energy_intensity_level_final_energy_megajoules_per_usd_2005_ppp,
    primary_energy_ppp_mj = energy_intensity_level_primary_energy_megajoules_per_usd_2005_ppp,
    energy_agri_mj = energy_intensity_agricultural_sector_megajoules_per_usd_2005,
    energy_industrial_mj = energy_intensity_industrial_sector_megajoules_per_usd_2005,
    energy_others_mj = energy_intensity_other_sectors_megajoules_per_usd_2005,
    primary_energy_savings_tj = energy_savings_primary_energy_terajoules,
    fp_ratio_pct = final_to_primary_energy_ratio_pct,
    # renewable
    renewable_output_pct = perc_renewable_of_total_electricity_output,
    renewable_gwh = renewable_energy_electricity_output_gigawatt_hours,
    renewable_capacity_gw = renewable_energy_installed_capacity_gigawatts,
    renewable_capacity_pct = share_of_renewable_capacity_in_total_capacity_pct,
    # others
    thermal_eff_pct = thermal_efficiency_in_power_supply_pct,
    total_output_gwh = total_electricity_output_gigawatt_hours,
    total_final_consumed_tj = total_final_consumption_terajoules,
    total_final_eu = total_final_energy_consumption_tfec,
    total_capacity_gw = total_installed_generation_capacity_gigawatts,
    total_primary_tj = total_primary_energy_supply_terajoules,
    power_lost_pct = transmission_and_distribution_losses_pct
  ) %>%
  rename_with(~ gsub("_consumption_tfec_pct", "_eu_pct", .x)) %>%
  rename_with(~ gsub("_consumption_terajoules", "_consumed_tj", .x))
```

For further analysis, let's also aggregate the countries by region.

``` {r}
energy <- energy %>%
  mutate(region = countrycode(country, origin = "country.name", destination = "region"))

unique(energy$region)

energy <- energy %>%
  mutate(region = ifelse(is.na(region) & country == "Mayotte", "Sub-Saharan Africa", region)) %>%
  mutate(region = ifelse(is.na(region) & country == "Puerto Rico", "Latin America & Caribbean", region)) %>%
  mutate(region = ifelse(is.na(region) & country == "Reunion", "Sub-Saharan Africa", region))
```

We can now inspect for N/A values. Using the naniar package, we can get the percentage of N/A values for each variable and visualize their distribution. A quick check also lets us know that no row is completely filled with N/A values. 

We can note that several energy types have a large amount of N/A values, namely geothermal energy, liquid biofuels, and wind energy. These values must be properly imputed before further exploratory data analysis.

``` {r}
energy %>%
  miss_var_summary() %>%
  slice_head(n = 20)
vis_miss(energy) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
sum(rowSums(is.na(energy)) == ncol(energy))
```

### Fuel and Electricity Access

Despite the high percentage of missing values for variables regarding fuel/electricity access, these variables will not be dropped as their missingness is not random; access is measured every 10 years. We can instead store access into a separate dataframe for analysis.

``` {r}
access <- energy %>%
  select(country, country_code, year, rural_access_fuel, urban_access_fuel, total_access_fuel, rural_access_elec, urban_access_elec, total_access_elec, region)
access <- access %>%
  filter(!(year > 1990 & year < 2000) & !(year > 2000 & year < 2010))
access
```

We shall first check for the top countries with access to non-solid fuel and electricity. In reading through the data would show us that it's recurring for nations to be at 100% electricity access, while fuel_access is at 95% for the

``` {r}
highest_access_countries_fuel <- access %>%
  group_by(country) %>%
  filter(min(total_access_fuel, na.rm = TRUE) == 95) %>%
  distinct(country)

highest_access_countries_fuel

perfect_access_countries_elec <- access %>%
  group_by(country) %>%
  filter(min(total_access_elec, na.rm = TRUE) == 100) %>%
  distinct(country)

perfect_access_countries_elec
```

We can observe that there are a lot of nations with near-perfect to perfect access to fuel and electricity over the last 20 years due to rapid urbanization, small land area, and advanced economy; it must be noted though that this does not mean a perfect distribution of electricity. Therefore, we shall instead focus on which countries continue to have the *least* access to fuel and electricity.

``` {r}
lowest_access_countries_fuel <- access %>%
  group_by(year) %>%
  mutate(fuel_rank = min_rank(total_access_fuel)) %>%
  arrange(year, fuel_rank) %>%
  slice_min(order_by = fuel_rank, n = 10) %>%
  ungroup() %>%
  distinct(country)

print(lowest_access_countries_fuel)

lowest_access_countries_elec <- access %>%
  group_by(year) %>%
  mutate(elec_rank = min_rank(total_access_elec)) %>%
  arrange(year, elec_rank) %>%
  slice_min(order_by = elec_rank, n = 10) %>%
  ungroup() %>%
  distinct(country)

print(lowest_access_countries_elec)
```

We can now map out these values.

``` {r}
world_map <- map_data("world")

# highest

highest_fuel_countries <- unique(highest_access_countries_fuel$country)
highest_elec_countries <- unique(perfect_access_countries_elec$country)
world_map_with_regions <- map_data("world") %>%
  mutate(world_region = countrycode(region, origin = "country.name", destination = "region"))

highlighted_map_highest <- world_map %>%
  mutate(category = case_when(
    region %in% highest_fuel_countries & region %in% highest_elec_countries ~ "Highest Access to Both",
    region %in% highest_fuel_countries ~ "Highest Non-solid Fuel Access",
    region %in% highest_elec_countries ~ "Highest Electricity Access",
    TRUE ~ "None"
  )) %>%
  filter(category != "None")

highlighted_map_highest <- world_map %>%
  mutate(category = case_when(
    region %in% highest_fuel_countries & region %in% highest_elec_countries ~ "Highest Access to Both",
    region %in% highest_fuel_countries ~ "Highest Non-solid Fuel Access",
    region %in% highest_elec_countries ~ "Highest Electricity Access",
    TRUE ~ "None"
  )) %>%
  filter(category != "None")

ggplot() +
  geom_polygon(data = world_map_with_regions,
               aes(x = long, y = lat, group = group, fill = world_region),
               color = "#ffffff", linewidth = 0.1) +
  scale_fill_viridis_d(option = "magma", alpha = 0.2, name = "World Regions") +
  
  new_scale_fill() +
  
  geom_polygon(data = highlighted_map_highest, 
               aes(x = long, y = lat, group = group, fill = category)) +
  geom_polygon(data = highlighted_map_highest, 
               aes(x = long, y = lat, group = group, fill = category),
               color = "white", linewidth = 0.05) +
  scale_fill_brewer(palette = "Dark2", name = "Access") +
  labs(title = str_wrap("Countries with Highest Access to Non-solid Fuel and Electricity", width = 40), x= "Longitude", y = "Latitude") +
  
  coord_quickmap()

# lowest

lowest_fuel_countries <- unique(lowest_access_countries_fuel$country)
lowest_elec_countries <- unique(lowest_access_countries_elec$country)

highlighted_map_lowest <- world_map %>%
  mutate(category = case_when(
    region %in% lowest_fuel_countries & region %in% lowest_elec_countries ~ "Lowest Access to Both",
    region %in% lowest_fuel_countries ~ "Lowest Non-solid Fuel Access",
    region %in% lowest_elec_countries ~ "Lowest Electricity Access",
    TRUE ~ "None"
  )) %>%
  filter(category != "None")

ggplot() +
  geom_polygon(data = world_map_with_regions,
               aes(x = long, y = lat, group = group, fill = world_region),
               color = "#ffffff", linewidth = 0.1) +
  scale_fill_viridis_d(option = "magma", alpha = 0.2, name = "World Regions") +
  
  new_scale_fill() +
  
  geom_polygon(data = highlighted_map_lowest, 
               aes(x = long, y = lat, group = group, fill = category)) +
  geom_polygon(data = highlighted_map_lowest, 
               aes(x = long, y = lat, group = group, fill = category),
               color = "white", linewidth = 0.05) +
  scale_fill_brewer(palette = "Dark2", name = "Access") +
  labs(title = str_wrap("Countries with Lowest Access to Non-solid Fuel and Electricity", width = 40), x = "Longitude", y = "Latitude") +
  
  coord_quickmap()
```

We notice that countries with the highest access are located around North America, Europe & Central Asia, and East Asia & Pacific, while those with the lowest access are generally in Sub-Saharan Africa and South Asia.

### LMDI Indices

``` {r}
geom_mean <- function(x) {
  exp(mean(log(x[x > 0 & !is.na(x)])))
}

regional_tidy <- energy %>%
  filter(!is.na(activity_index)) %>%
  group_by(region, year) %>%
  summarise(
    Activity = geom_mean(activity_index),
    Efficiency = geom_mean(efficiency_index),
    Structure = geom_mean(structure_index),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(Activity, Efficiency, Structure),
    names_to = "effect",
    values_to = "index_value"
  )

ggplot(regional_tidy, aes(x = year, y = index_value, color = effect, group = effect)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 1) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
  facet_wrap(~ region) +
  scale_color_manual(
    values = c(
      "Activity"   = "#FB8861",
      "Efficiency" = "#B63679",
      "Structure"  = "#51127C"
    )
  ) +
  labs(
    title = "Regional Multiplicative LMDI Drivers Over Time",
    subtitle = "Aggregated using geometric means (Base Year = 1.0)",
    x = "Year",
    y = "Index Value",
    color = "Decomposition Factor"
  ) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold", size = 10)
  )
```

From the graph above, the first most noticeable pattern is the sudden drop as we approach the 2010s. This is due to the 2008 financial crisis, which largely affected economic growth across all nations. It can be observed, however, some regions have steeper economic growth over time, such as East Asia & Pacific and South Asia. Others, such as North America and Europe & Central Asia have steadier growth. 

The structure index also remains close to the baseline of 1.0, meaning that across the world, changes in economic sectors have a negligible effect upon energy usage.

For efficiency, its index goes downwards for almost all regions, except Latin America & Caribbean and Middle East & North Africa, meaning that energy usage was not offset much by developments in technology. It can be noted for these regions that technological efficiency did not combat activity growth as efficiently as other nations. 

### Renewable Energy Distribution

``` {r}
energy_absolute_long <- energy %>%
  select(
    region, country, year,
    Biogas = biogas_consumed_tj,
    Geothermal = geothermal_energy_consumed_tj,
    Hydro = hydro_energy_consumed_tj,
    `Liquid Biofuels` = liquid_biofuels_consumed_tj,
    Marine = marine_consumed_tj,
    `Modern Biomass` = modern_biomass_consumed_tj,
    Solar = solar_energy_consumed_tj,
    `Traditional Biomass` = traditional_biomass_consumed_tj,
    Waste = waste_energy_consumed_tj,
    Wind = wind_energy_consumed_tj
  ) %>%
  pivot_longer(
    cols = Biogas:Wind,
    names_to = "renewable_type",
    values_to = "consumption_tj"
  ) %>%
  mutate(consumption_tj = replace_na(consumption_tj, 0))

absolute_regional_summary <- energy_absolute_long %>%
  group_by(region, renewable_type) %>%
  summarise(total_tj = sum(consumption_tj, na.rm = TRUE), .groups = "drop")

print(absolute_regional_summary)

ggplot(absolute_regional_summary, aes(x = region, y = total_tj, fill = renewable_type)) +
  geom_bar(stat = "identity", position = "fill") +
  theme_minimal() +
  labs(
    title = "Proportional Composition of Absolute End-User Consumption",
    x = "Region",
    y = "Proportion of Total Volumetric Mix",
    fill = "Renewable Type"
  ) +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

yearly_absolute_trend <- energy_absolute_long %>%
  group_by(region, year, renewable_type) %>%
  summarise(total_tj = sum(consumption_tj, na.rm = TRUE), .groups = "drop")

ggplot(yearly_absolute_trend, aes(x = year, y = total_tj, color = renewable_type)) +
  geom_line(size = 0.7) +
  facet_wrap(~region, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Absolute End-User Consumption Volumetric Trends Over Time",
    x = "Year",
    y = "Total Consumption (Terajoules)",
    color = "Renewable Type"
  )
```

``` {r}
top_5_countries <- energy %>%
  rowwise() %>%
  mutate(total_country_consumption = sum(c_across(ends_with("_consumed_tj")), na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(country) %>%
  summarise(cumulative_consumption = sum(total_country_consumption, na.rm = TRUE)) %>%
  arrange(desc(cumulative_consumption)) %>%
  slice_max(order_by = cumulative_consumption, n = 5) %>%
  pull(country)

energy_top5_comparison <- energy %>%
  filter(country %in% top_5_countries) %>%
  select(
    country, region, year, total_capacity_gw,
    Biogas = biogas_consumed_tj,
    Geothermal = geothermal_energy_consumed_tj,
    Hydro = hydro_energy_consumed_tj,
    `Liquid Biofuels` = liquid_biofuels_consumed_tj,
    Marine = marine_consumed_tj,
    `Modern Biomass` = modern_biomass_consumed_tj,
    Solar = solar_energy_consumed_tj,
    `Traditional Biomass` = traditional_biomass_consumed_tj,
    Waste = waste_energy_consumed_tj,
    Wind = wind_energy_consumed_tj
  ) %>%
  pivot_longer(
    cols = Biogas:Wind,
    names_to = "renewable_type",
    values_to = "consumption_tj"
  ) %>%
  mutate(
    consumption_tj = replace_na(consumption_tj, 0),
    total_capacity_gw = replace_na(total_capacity_gw, 0)
  )

top5_summary <- energy_top5_comparison %>%
  group_by(country) %>%
  summarise(
    total_consumed_tj = sum(consumption_tj, na.rm = TRUE),
    average_capacity_gw = mean(total_capacity_gw, na.rm = TRUE)
  ) %>%
  arrange(desc(total_consumed_tj))

print(top5_summary)
```

``` {r}
plot_data_consumption <- energy_top5_comparison %>%
  group_by(country, renewable_type) %>%
  summarise(Value = sum(consumption_tj, na.rm = TRUE), .groups = "drop") %>%
  mutate(Metric = "End-User Consumption (TJ)")

plot_data_capacity <- energy_top5_comparison %>%
  group_by(country) %>%
  summarise(Value = mean(total_capacity_gw, na.rm = TRUE), .groups = "drop") %>%
  mutate(renewable_type = "Total Infrastructure", Metric = "Avg Infrastructure Capacity (GW)")

combined_plot_data <- bind_rows(plot_data_consumption, plot_data_capacity)

# Plot using facets to separate the incompatible units (TJ vs GW)
ggplot(combined_plot_data, aes(x = country, y = Value, fill = renewable_type)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~Metric, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Top 5 Countries: Renewable Consumption vs. Infrastructure Capacity",
    x = "Country",
    y = "Scale Value",
    fill = "Energy Component"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

yearly_totals_top5 <- energy_top5_comparison %>%
  group_by(country, year, total_capacity_gw) %>%
  summarise(yearly_consumption_tj = sum(consumption_tj, na.rm = TRUE), .groups = "drop")

# Scatter plot showing historical trajectory paths
ggplot(yearly_totals_top5, aes(x = total_capacity_gw, y = yearly_consumption_tj, color = country)) +
  geom_point(size = 2, alpha = 0.6) +
  geom_path(aes(group = country), arrow = arrow(length = unit(0.15, "cm"))) + # Shows direction over time
  theme_minimal() +
  labs(
    title = "Historical Trajectory: Infrastructure Capacity vs. Actual Consumption",
    subtitle = "Arrows trace chronological growth over the years",
    x = "Total Infrastructure Capacity (GW)",
    y = "Total End-User Consumption (Terajoules)",
    color = "Country"
  )
```

``` {r}
regional_top_countries <- energy %>%
  # Filter out unwanted regions
  # Sum up all absolute end-user consumption columns for every individual row
  rowwise() %>%
  mutate(total_row_consumption = sum(c_across(ends_with("_consumed_tj")), na.rm = TRUE)) %>%
  ungroup() %>%
  # Group by region and country to get lifetime totals across all available years
  group_by(region, country) %>%
  summarise(cumulative_consumption = sum(total_row_consumption, na.rm = TRUE), .groups = "drop_last") %>%
  # Rank countries locally inside their respective region profiles
  mutate(consumption_rank = min_rank(desc(cumulative_consumption))) %>%
  # Keep only the top 5 positions for each region group
  filter(consumption_rank <= 5) %>%
  ungroup()

# Step 2: Filter your core dataset to keep ONLY these regional leaders
energy_regional_top5 <- energy %>%
  filter(country %in% regional_top_countries$country) %>%
  select(
    country, region, year, total_capacity_gw,
    Biogas = biogas_consumed_tj,
    Geothermal = geothermal_energy_consumed_tj,
    Hydro = hydro_energy_consumed_tj,
    `Liquid Biofuels` = liquid_biofuels_consumed_tj,
    Marine = marine_consumed_tj,
    `Modern Biomass` = modern_biomass_consumed_tj,
    Solar = solar_energy_consumed_tj,
    `Traditional Biomass` = traditional_biomass_consumed_tj,
    Waste = waste_energy_consumed_tj,
    Wind = wind_energy_consumed_tj
  ) %>%
  pivot_longer(
    cols = Biogas:Wind,
    names_to = "renewable_type",
    values_to = "consumption_tj"
  ) %>%
  mutate(
    consumption_tj = replace_na(consumption_tj, 0),
    total_capacity_gw = replace_na(total_capacity_gw, 0)
  )

viz_consumption <- energy_regional_top5 %>%
  group_by(region, country, renewable_type) %>%
  summarise(total_tj = sum(consumption_tj, na.rm = TRUE), .groups = "drop")

ggplot(viz_consumption, aes(x = country, y = total_tj, fill = renewable_type)) +
  geom_bar(stat = "identity", position = "stack") +
  # Create separate plot windows for each region with custom adjusted Y-axes scales
  facet_wrap(~region, scales = "free") + 
  theme_minimal() +
  labs(
    title = "Top 5 End-User Energy Consuming Countries per Region",
    x = "Country",
    y = "Total Energy Consumption (Terajoules)",
    fill = "Renewable Type"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    strip.text = element_text(face = "bold", size = 10)
  )

# Aggregate maximum grid capacity metrics for visualization
viz_capacity <- energy_regional_top5 %>%
  group_by(region, country) %>%
  summarise(avg_capacity = mean(total_capacity_gw, na.rm = TRUE), .groups = "drop")

ggplot(viz_capacity, aes(x = country, y = avg_capacity, fill = region)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_wrap(~region, scales = "free") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Infrastructure Grid Footprint for Top Regional Energy Users",
    x = "Country",
    y = "Average Generation Capacity (Gigawatts)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
    strip.text = element_text(face = "bold", size = 10)
  )
```